---
title: "Project_part2"
author: "Himanshu Raj"
date: "11 May 2018"
output:
  html_document:
    keep_md: yes
  pdf_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Project Part 2: Basic Inferential Data Analysis

#### Overview
In the second portion of the project, we're going to analyze the ToothGrowth data in the R datasets package. We have to perform basic exploratory data analysis and summary of data.

#### Loading Libraries and data
```{r lib}
library(ggplot2)
data(ToothGrowth)
attach(ToothGrowth)
```

#### About Data
The data talks about the Effect of Vitamin C on Tooth Growth in Guinea Pigs

Description:
The response is the length of odontoblasts (cells responsible for tooth growth) in 60 guinea pigs. Each animal received one of three dose levels of vitamin C (0.5, 1, and 2 mg/day) by one of two delivery methods, (orange juice or ascorbic acid (a form of vitamin C and coded as VC).


#### Loking at the basic stats
```{r stat}
str(ToothGrowth)
```


#### Exploratory Data Analysis

We will be plotting a scatterplot to demonstrate the relationship between Tooth Lengths with respect to Dose Levels and Delivery Methods

```{r dataVis}
ToothGrowth$dose<- as.factor(ToothGrowth$dose)
ggplot(ToothGrowth, aes(x= dose, y= len, color= supp)) + geom_boxplot() + xlab("Dose of Vitamin C(mg/day)") + ylab("Length of odontoblasts")
```

```{r dataVis2}
ggplot(ToothGrowth, aes(x= len, color= dose)) + geom_density(alpha=0.5) + coord_flip() + xlab("Length of odontoblasts") + ylab("Density") + scale_fill_manual(values=c("orange","purple"))
```


#### Loking at the data summary 
```{r summry}
summary(ToothGrowth)
```

#### Using hypothesis tests to compare tooth growth by supp and dose.

##### Using Supplement Delivery Method As A Factor:

Null Hypothesis: here is no correlation between the Delivery Method and Tooth Length.

```{r test1}
t.test(len ~ supp, paired = F, var.equal = F, data = ToothGrowth)
```

As per the above test we gather the inference that the 95% confidence interval is `[-0.1710156, 7.5710156]` and the p-value is `0.06063`(6.06%), which is greater than `0.05`(5%). Hence, the NULL Hypothesis cannont be rejected.

Therfore, from the above statements and inferences we conclude that there is no correlation between Delivery Method and Tooth Length.

##### Using Supplement Dosage Level As A Factor:

Dividing the data based on the dose level:
```{r divide}
Dose_1 <- subset(ToothGrowth, dose %in% c(0.5, 1.0))
Dose_2 <- subset(ToothGrowth, dose %in% c(0.5, 2.0))
Dose_3 <- subset(ToothGrowth, dose %in% c(1.0, 2.0))
```

Analyzing the data for correlation between the Dose Level and change in Tooth Growth, assuming unequal variances between the two groups.

NULL Hypothesis: For the following three t-tests, there is no correlation between the Dose Level and Tooth Length.

Testing for Dose level in 0.5 and 1.0
```{r test2_1}
t.test(len ~ dose, paired = F, var.equal = F, data = Dose_1)
```

Here, the 95% confidence interval is `[-11.983781, -6.276219]`, which does not contain zero and the p-value is `1.268e-07`, which is less than 0.05. Hence, we can safely reject the NULL Hypothesis.

Testing for Dose level in 0.5 and 2.0
```{r test2_2}
t.test(len ~ dose, paired = F, var.equal = F, data = Dose_2)
```

Here, the 95% confidence interval is `[-18.15617, -12.83383]`, which does not contain zero and the p-value is `4.398e-14`, which is less than `0.05`. Hence, we can safely reject the NULL Hypothesis.

Testing for Dose level in 1.0 and 2.0
```{r test2_3}
t.test(len ~ dose, paired = F, var.equal = F, data = Dose_3)
```

Here, the 95% confidence interval is `[-8.996481, -3.733519]`, which does not contain zero and the p-value is `1.906e-05`, which is less than `0.05`. Hence, we can safely reject the NULL Hypothesis.

From these three t-tests, we conclude that, There is significant correlation between the Dose Level and Tooth Length.

#### Conclusions:
Increase in Supplement Dose Levels leads to overall increase in Tooth Length.

