---
title: "Statistical_inference_Course_project"
author: "Himanshu Raj"
date: "9 May 2018"
output:
  word_document: default
  pdf_document: default
  html_document:
    keep_md: yes
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Project Part 1: A simulation exercise

#### Overview

In this project we will investigate the exponential distribution in R and compare it with the Central Limit Theorem. The exponential distribution can be simulated in R with rexp(n, lambda) where lambda is the rate parameter.

#### Starting with the Simulation
The mean of the Exponential distribution is $\frac{1}{\lambda}$ and the standard deviation is also $\frac{1}{\lambda}$. We will investigate the distribution of averages of 40 exponentials.
T The exponential distribution can be simulated in R with `rexp(n, lambda)` where lambda is the rate parameter. We are setting `lambda = 0.2` for all of the simulations.

#### Loading Libraries

```{r lib}
library(ggplot2)
```

Initializing the simulation controlling variables.
```{r init}
noSim <- 1000
sampSize <- 40
lambda <- 0.2
```

Setting the seed of the Random Number Generator and creating matrix for 1000 rows and 40 samples
```{r seed}
set.seed(123)
simMatrix <- matrix(rexp(n = noSim * sampSize, rate = lambda), noSim, sampSize)
```

Creating a vector of thousand rows containing the mean of each row of the simMatrix and combining into dataframe
```{r mean}
simMean <- rowMeans(simMatrix)
simData <- data.frame(cbind(simMatrix, simMean))
dim(simData)
```

### Plotting and Visualising the data
```{r plot}
ggplot(data = simData, aes(simData$simMean)) + 
  geom_histogram(breaks = seq(2, 9, by = 0.2), col = "blue", aes(fill = ..count..)) + 
  labs(title = "Histogram of Mean Distribution", x = "Simulation Means", y = "Frequency") + 
  geom_vline(aes(xintercept=mean(simData$simMean)), color="red", 
             linetype="dashed", size=1)
```

### Comparing Sample mean to the theoretical mean of the distribution.

```{r comMean}
actualMean <- mean(simMean)
theoreticalMean <- (1 / lambda)
```

The actual mean of the simulated mean sample data is `r actualMean` and the theoretical mean is `r theoreticalMean`.

Thus, we can see that the actual mean of the simulated mean sample data is very close to the theoretical mean of original data distribution.

### Comparing the sample variance and the theoretical variance of the distribution.

```{r comVar}
actualVariance <- var(simMean)
theoreticalVariance <- ((1 / lambda) ^ 2) / sampSize
```

The actual variance of the simulated mean sample data is `r actualVariance` and the  theoretical variance is `r theoreticalVariance`.

Thus, we can see that the actual variance of the simulated mean sample data is very close to the theoretical variance of original data distribution.

### Observing the distribution

To prove that the simulated mean sample data approximately follows the Normal distribution, we perform the following three steps:

Step 1: Create an approximate normal distribution and see how the sample data alligns with it.

```{r Distrib}
qplot(simMean, geom = 'blank') + 
  geom_line(aes(y=..density.., colour='Empirical'), stat='density', size=1) + 
  stat_function(fun=dnorm, args=list(mean=(1/lambda), sd=((1/lambda)/sqrt(sampSize))), 
                aes(colour='Normal'), size=1) + 
  geom_histogram(aes(y=..density.., fill=..density..), alpha=0.4, 
                 breaks = seq(2, 9, by = 0.2), col='red') + 
  scale_fill_gradient("Density", low = "green", high = "blue") + 
  scale_color_manual(name='Density', values=c('brown', 'blue')) + 
  theme(legend.position = c(0.85, 0.60)) + 
  labs(title = "Mean Density Distribution", x = "Simulation Means", y = "Density")
```  
  
From above histogram, the simulated mean sample data can be adequately approximated with the normal distribution.

Step 2: Compare the 95% confidence intervals of the simulated mean sample data and the theoretical normally distributed data.

```{r Distrib2}
actualConfInterval <- actualMean+c(-1,1)*1.96*sqrt(actualVariance)/sqrt(sampSize)
theoreticalConfInterval <- theoreticalMean+c(-1,1)*1.96*
  sqrt(theoreticalVariance)/sqrt(sampSize)
```

Actual 95% confidence interval is [4.7414712, 5.2317681] and Theoretical 95% confidence interval is [4.755, 5.245] and we see that both of them are approximately same.

Step 3: q-q Plot for Qunatiles.
```{r Distrib3}
qqnorm(simMean)
qqline(simMean)
```

The actual quantiles also closely match the theoretical quantiles, hence the above three steps prove that the distribution is approximately normal.
